The Role of Advection Versus Diffusion in Intraventricular Transport

### 2025-02-22 paper/main.tex
updates discussion

\title{Advection Versus Diffusion in Intraventricular Transport}

@@

% Author: Please give full first and last names for authors and include * after the name of all corresponding authors

@@

% Affiliations: Please provide adacemic titles (Prof. or Dr.) for all authors where applicable, and include an institutional email address for all corresponding authors
\begin{affiliations}
H.~Herlyng, Dr.~A.~J.~Ellingsrud, Dr.~M.~Kuchta, Dr.~M.~E.~Rognes\\
Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Kristian Augusts gate 23, 0164 Oslo, Norway\\
Email Address: hherlyng@simula.no

Dr.~M.~E.~Rognes \\
K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway
\\
Dr.~I.~Jeong, Dr.~N.~Jurisch-Yaksi\\
Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Erling Skjalgsons gate 1, 7491 Trondheim, Norway \\
Email Address: nathalie.jurisch-yaksi@ntnu.no

@@

\keywords{cerebrospinal fluid, brain ventricles, fluid flow, transport}

@@

% Experimental section
\section{Computational Methods}
\subsection{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a
3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Crefpanels{fig:fig1}{d}).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The \emph{standard} mesh consisted of 132 134 tetrahedral cells and had a
maximal (minimal) edge length of 14.9 \textmu m (4.68 \textmu m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles, which are connected by ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia (\Crefpanels{fig:fig1}{d}, magenta markers)~\cite{Olstad2019CiliaryDevelopment}.
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b.} Particle tracking data visualizing flow directionality in the middle ventricle.
    \textbf{c.} Flow velocity magnitudes in the middle ventricle from particle tracking data.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Streamlines simulated with the baseline flow model at
    a time 25\% into the cardiac cycle. %t=0.112613 s
    \textbf{h.} Streamlines simulated with the baseline flow model at
    a time 50\% into the cardiac cycle. %t=0.225225 s
    \textbf{i.} Streamlines simulated with the baseline flow model at
    a time 75\% into the cardiac cycle. %t=0.315315 s
    \textbf{j.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of panels \textbf{f, h--j} are indicated.
    }
    \label{fig:fig1}
\end{figure}

The morphology and geometry of the brain ventricles vary across zebrafish
individuals, and under physiological and pathological conditions. 
Motivated by an interest in how ventricular geometry impacts solute distribution,
we also consider four variations in the geometry \fixme{(cross-sectional area
reduction in parentheses):
by shrinking the fore-mid brain connection (66 \%), 
the middle ventricle (33 \%), the mid-hind brain connection (43 \%), and all three simultaneously.} 
We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.
% Cross-sectional areas of ducts normal to flow direction
% Modified region       Original mesh        Deformed mesh
% Fore                  3.36e-3 mm^2         1.15e-3 mm^2
% Mid                   6.11e-3 mm^2         4.08e-3 mm^2
% Hind                  2.91e-3 mm^2         1.67e-3 mm^2

\subsection{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.70 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a force $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential forces $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\mathbf{J}$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\Cref{eq:stokes_eqs},
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gamma_p$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling the Dendra2 Fluorescent Protein (Dendra2)
used in this study, the Starmaker+Green Fluorescent Protein (STM-GFP) reported in
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and exosomes.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation:
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
solvent (in this case the CSF), and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with $R=200$ nm. This radius covers the slowest diffusing exosomes,
since the smaller exosomes would have diffusion coefficients similar to the other molecules.
For STM-GFP, we estimated the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients used in the simulations
are $D_1$ (exosomes), $D_2$ (STM-GFP) and $D_3$ (Dendra2)~(\Cref{tab:diff_coeff_table}).

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight}
        & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\Cref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
\end{table}

@@

%-----DISCRETIZATION AND MODEL IMPLEMENTATION-----%
\subsection{Numerical approximation of the Stokes equations}
The Stokes equations~\Cref{eq:stokes_eqs} are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization.
We set the timestep size $\Delta t = 0.02252$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM) elements for the velocity.
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements~\cite{Brezzi1985TwoProblems},
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 943 820.
\fixme{The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
To ensure stability of the scheme, continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.}
Since the divergence of the velocity is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly on each
element~\cite{Boffi2008FiniteProblem}, ensuring mass conservation.

@@

\subsection{Generation of transgenic line to study CSF \fixme{and transport} dynamics} %\emph{Tg(5xuas:Signal-Dendra2)\textsuperscript{nw21Tg}}
To generate a transgenic line expressing the photoconvertible protein Dendra2 in
the brain ventricles, we fused the open reading frame (ORF)
of the neuropeptide y (npy) signal peptide
(\emph{npy}: ENSDARG00000036222, Q1LW93) to the N-terminal of
zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and
3’ ends (GenScript Biotech) and inserted into the
pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish}
through restriction enzyme cloning (GenScript Biotech)
to generate the 5xuas:Signal-Dendra2 plasmid DNA.

@@

\section{Discussion}
We have presented a computational model of CSF flow and solute transport within the brain ventricles.
In this study, we used the larval zebrafish brain ventricles since its geometry,
flow components and kinetics are well described, thereby allowing us to validate
our simulation experimentally. Nevertheless, our simulations can be implemented in
other systems in the future when their flow parameters are comprehensively resolved.  
By imposing tangential traction and normal pressure on the CSF flow, we were able to reproduce
flow fields with similar features as observed experimentally, consisting of vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
These results compared well with experimental data performed in vivo in zebrafish.
Then, by using both a computational model and zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role in protein transport.
Finally, we reported how local cilia paralysis and ventricular geometries
affect solute distribution. Altogether, we present not only new findings on the role of
diffusion and advection, but also provide a novel framework to
study CSF flow and transport in complex geometries.

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the duct volumes impacts
solute distribution and the geometry used for the model was larger than in experiments.
Alternatively, it may relate to the way we defined the photoconversion site (ROI 1)
in the simulations, since the volume of the photoconversion region determines
the total amount of solute, and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24h old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles and render the dynamics slower.

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 200 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 850 s), where transport is largely driven by diffusion.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=1.41 \ \mu$m/s over the same distance would happen
on a timescale of the order of $t_a=L/\overline{U}=141$ s. Comparing this with the minuscule values of
$\cbar_6$ after simulating transport with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

>>>>

\title{The Role of Advection Versus Diffusion in Intraventricular Transport}

@@

% Author: Please give full first and last names for authors and
% include * after the name of all corresponding authors

@@

% Affiliations: Please provide adacemic titles (Prof. or Dr.) for all authors where applicable,
% and include an institutional email address for all corresponding authors
\begin{affiliations}
H.~Herlyng, Dr.~A.~J.~Ellingsrud, Dr.~M.~Kuchta, Dr.~M.~E.~Rognes\\
Department of Numerical Analysis and Scientific Computing,
Simula Research Laboratory, Kristian Augusts gate 23, 0164 Oslo, Norway\\
Email Address: hherlyng@simula.no
\\
Dr.~M.~E.~Rognes\\
K.~G.~Jebsen Center for Brain Fluid Research, Oslo, Norway
\\
Dr.~I.~Jeong, Dr.~N.~Jurisch-Yaksi\\
Department of Clinical and Molecular Medicine,
Norwegian University of Science and Technology, Erling Skjalgsons gate 1, 7491 Trondheim, Norway \\
Email Address: nathalie.jurisch-yaksi@ntnu.no

@@

\keywords{cerebrospinal fluid, brain ventricles, transport, finite element method}

@@

% Experimental section
\section{Computational Methods}
\subsection{Image-based computational geometries of zebrafish brain ventricles}
To model flow and transport in zebrafish larval brain ventricles, we generated a
3D representation of the brain ventricular walls using confocal imaging data of a zebrafish embryo
injected intraventricularly with a 70 kDa dye at 2 dpf~\cite{Olstad2019CiliaryDevelopment}. 
From the imaging data, we constructed a computational mesh of the ventricles (\Crefpanels{fig:fig1}{d}).
The interior volume was meshed with fTetWild~\cite{Hu2020FastWild}. 
The standard mesh consisted of 132 134 tetrahedral cells and had a
maximal (minimal) edge length of 14.9 \textmu m (4.68 \textmu m). 
At 2 dpf, the ventricular system consists mainly of 3 cavities: the telencephalic (anterior),
the di-/mesencephalic (middle) and the rhombencephalic (posterior) ventricles,
which are connected by interventricular ducts. 
Parts of the ventricular surfaces were marked to be lined with
motile cilia (\Crefpanels{fig:fig1}{d}, magenta markers)~\cite{Olstad2019CiliaryDevelopment}.
\begin{figure}%[hbt!]
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b.} Particle tracking data visualizing flow directionality in the middle ventricle.
    \textbf{c.} Flow velocity magnitudes in the middle ventricle from particle tracking data.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Streamlines simulated with the baseline flow model at
    a time 25\% into the cardiac cycle. %t=0.112613 s
    \textbf{h.} Streamlines simulated with the baseline flow model at
    a time 50\% into the cardiac cycle. %t=0.225225 s
    \textbf{i.} Streamlines simulated with the baseline flow model at
    a time 75\% into the cardiac cycle. %t=0.315315 s
    \textbf{j.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of panels \textbf{f, h--j} are indicated.
    }
    \label{fig:fig1}
\end{figure}

The morphology and geometry of the brain ventricles vary across zebrafish
individuals, and under physiological and pathological conditions~\cite{Olstad2019CiliaryDevelopment}. 
Motivated by an interest in how ventricular geometry impacts solute distribution,
we also consider four variations in the geometry (cross-sectional area
reduction in parentheses):
by shrinking the fore-mid brain connection (66 \%), 
the middle ventricle (33 \%), the mid-hind brain connection (43 \%), and all three simultaneously.
We used Blender~\cite{Community2018BlenderPackage} to modify the brain ventricles geometry.
% Cross-sectional areas of ducts normal to flow direction
% Modified region       Original mesh        Deformed mesh
% Fore                  3.36e-3 mm^2         1.15e-3 mm^2
% Mid                   6.11e-3 mm^2         4.08e-3 mm^2
% Hind                  2.91e-3 mm^2         1.67e-3 mm^2

\subsection{Computational CSF dynamics}
The beating motion of the cilia generates CSF flow with steady rotational
structures~\cite{Olstad2019CiliaryDevelopment}
within the brain ventricles (\Crefpanels{fig:fig1}{a--c}).
In addition, cardiac pulsations induce pulsatile CSF flow.
We model this flow of CSF in the brain ventricles by the time-dependent,
incompressible Stokes equations which read as follows.
Find the CSF velocity $\uu = \uu(\xx, t) = (u_x(\xx, t), u_y(\xx, t), u_z(\xx, t))$
and the CSF pressure $p = p(\xx, t)$ for $\xx \in \Omega$ and time $t>0$, such that
\begin{subequations}
    \begin{align}
      \rho \pdifft{\uu} - \nabla \cdot \bsig &= \mathbf{0}
      \quad \text{in } \Omega,\\
      \nabla \cdot \uu &= 0
      \quad  \text{in } \Omega,
  \end{align}
  \label{eq:stokes_eqs}% 
\end{subequations}%
where the stress tensor is defined as $\bsig(\uu, p) = 2\mu\beps(\uu) - p\mathbf{I}$
and $\beps(\uu) = \frac{1}{2}\left(\nabla \uu + (\nabla\uu)^T\right)$ is the strain-rate tensor. 
Bold-face characters denote vectors or tensors, and $\mathbf{I}$ is the identity tensor in three dimensions.
Furthermore, $\rho=1000 \ \mathrm{kg/m^3}$ is the CSF density
and $\mu=0.70 \ \mathrm{mPa\cdot s}$ is the dynamic viscosity
of the CSF~\cite{Bloomfield1998EffectsFluid}.
We introduce the tangential traction $\bsigpar$ as the
tangential component of the traction $\hat{\bsig}=\bsig\nn$:
\begin{equation*}
    \bsigpar = P_{\nn}(\hat{\bsig}) = (\mathbf{I} - \nn\otimes\nn)\hat{\bsig},
\end{equation*}
where $\nn$ is the outer unit normal of the boundary surface $\Gamma = \partial\Omega$,
and $P_{\nn}(\rr)$ is the tangential projection of a vector $\rr$ onto $\Gamma$.
We consider three types of boundary conditions:
(i) \emph{tangential traction} (no flow normal to the boundary
$\uu \cdot \nn = 0$ with a stress $\btau$ applied tangentially on the boundary $\bsigpar = \btau$),
(ii) \emph{normal pressure} (a pressure $\Tilde{p}$ applied in
the normal direction: $(\mu\nabla\uu - p\mathbf{I})\nn = \Tilde{p}(t)\nn$),
and (iii) \emph{free slip} (no flow normal to the boundary
$\uu \cdot \nn = 0$ and no tangential traction $\bsigpar = {\bf 0}$).
Unless noted otherwise, the free-slip condition is applied on the
boundary. We assume that the system starts at rest: $\uu(\xx, t=0) = \mathbf{0}$.

@@

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\Cref{eq:stokes_eqs},
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gamma_p$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling the Dendra2 Fluorescent Protein (Dendra2)
used in this study, the Starmaker+Green Fluorescent Protein (STM-GFP) reported in
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and exosomes.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
solvent (in this case the CSF), and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with $R=200$ nm. This radius covers the slowest diffusing exosomes,
a choice made because smaller exosomes would have diffusion coefficients
similar to the other molecules.
For STM-GFP, we estimated the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients used in the simulations
are $D_1$ (exosomes), $D_2$ (STM-GFP) and $D_3$ (Dendra2)~(\Cref{tab:diff_coeff_table}).

@@

\begin{table}[h!]
    \centering
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\Cref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight}
        & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 200 nm) & -- & $\num{1.63e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
\end{table}

@@

%-----DISCRETIZATION AND MODEL IMPLEMENTATION-----%
\subsection{Numerical approximation of the Stokes equations}
The Stokes equations~\Cref{eq:stokes_eqs} are discretized with a finite element method
in space, while in time, we employ a first-order implicit Euler discretization.
We set the timestep size $\Delta t = 0.02252$ s such that 20 timesteps represent one cardiac cycle.
For the spatial discretization, we use piecewise linear Brezzi-Douglas-Marini (BDM) elements for the velocity.
The BDM elements have degrees of freedom associated with integral moments of the
velocity component normal to the facets of elements~\cite{Brezzi1985TwoProblems},
thus we enforce the impermeability condition $\uu\cdot\nn=0$ strongly.
We impose the boundary conditions for both the tangential traction $\bsigpar$ and
the normal pressure $\Tilde{p}(t)$ weakly.
For the pressure, we use piecewise constant (zeroth order discontinuous Galerkin (DG)) elements.
The resulting total number of degrees of freedom at each time step is 943 820.
The element combination $\mathrm{BDM}_1$--$\mathrm{DG}_{0}$ yields a non-conforming
discretization scheme for the Stokes equations~\cite{Stenberg1989SomeEquations}.
To ensure stability of the scheme, continuity in tangential velocities on interior facets
is weakly enforced with a penalty parameter $\gamma=10$~\cite{Hong2016AEquations}.
Since the divergence of the velocity is a subspace of the pressure space for this scheme,
the divergence-free condition of the velocity is satisfied exactly on each
element, ensuring mass conservation.~\cite{Boffi2008FiniteProblem}

@@

\subsection{Generation of transgenic line to study CSF and transport dynamics} 
To generate a transgenic line expressing the photoconvertible protein Dendra2 in
the brain ventricles, we fused the open reading frame (ORF)
of the neuropeptide y (npy) signal peptide
(\emph{npy}: ENSDARG00000036222, Q1LW93) to the N-terminal of
zebrafish codon optimized Dendra2 DNA sequence.
The DNA sequence was synthesized with EcoR1 enzyme sites at the 5’ and
3’ ends (GenScript Biotech) and inserted into the
pT2MUASMCS vector~\cite{Asakawa2008GeneticZebrafish}
through restriction enzyme cloning (GenScript Biotech)
to generate the 5xuas:Signal-Dendra2 plasmid DNA.

@@

\section{Discussion}
In this study, we have presented a computational model of CSF flow and solute transport within the brain ventricles.
We used the larval zebrafish brain ventricles since its geometry,
flow components and kinetics are well described, thereby allowing us to validate
our simulation experimentally. Nevertheless, our computational framework can be adapted to and implemented in
other ventricular systems in the future, given that the flow parameters of such systems
are comprehensively resolved. By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

Using the simulated CSF flow, we modeled solute transport by an advection-diffusion equation.
The transport results compared well with experimental data performed in vivo in zebrafish.
Then, by using a computational model as well as zebrafish mutants with paralyzed cilia,
we identified that even though advection by cilia is dominant,
diffusion plays a significant role in protein transport.
Finally, we reported how local cilia paralysis and ventricular geometries
affect solute distribution. Altogether, we present not only new findings on the role of
diffusion and advection in intraventriuclar transport, but also provide a novel framework to
study CSF flow and transport in complex geometries.

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the interventricular ducts
impacts solute distribution, and the geometry used for the computational model was larger
than geometries used in experiments. Alternatively, dissimilarities in results may relate
to the way we defined the photoconversion site (ROI 1) in the simulations,
since the volume of the photoconversion region determines
the total amount of solute and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the time spans we consider
and render the dynamics slower.

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 200 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 850 s), a ventricle where transport is largely driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=1.41 \ \mu$m/s over the same
distance would happen on a timescale of the order of $t_a=L/\overline{U}=141$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.
